Data Analysis

### 2023-04-11 ms1_data_analysis.Rmd
Prediction plots for I, P, H

```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
# The source file loads the data preparation script, where the raw data is read and summarised for further analyses
source('code/log_ratio_extraction.R')
```

@@

Figure 1. Median, 1st, second quadrilles, 1.5 IRQ distance and outliers in biomass, richness diversity and density at each location/elevation, but some Yawan sites were extremely productive.

```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure S1 Characteistics"}

library(ggplot2)
library(dplyr)

@@

summary(trtlm)
summary(trtgls)

@@

Table S2. Two models testing difference of LRR of predator exclosures from zero at each elevation.

```{r}
sjPlot::tab_df(model_comp)
```

@@

      # np.test.p <- cor.test(lrrtd$x, lrrtd$y, method = c('pearson'))$p.value
      # Test for noramlity
      print(shapiro.test(lrrtd$x))
      print(shapiro.test(lrrtd$y))

@@

@@

ggplot(cpdata, aes(x = x, y = y)) +
  geom_point(size = 2, alpha = 0.4)+
  stat_smooth(data = cpdata %>%
                filter(sign == "signif"),
              method ="lm")+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  ylab("Effect size of insects")+
  xlab("Effect size")+
  theme_bw()+
  facet_grid(response*descriptor~site)

@@

ggplot(fidata, aes(x = x, y = y)) +
  geom_point(size = 2, alpha = 0.4)+
  stat_smooth(data = fidata %>%
                filter(sign == "signif"),
              method ="lm")+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  xlab("Predator removal")+
  ylab("Predator removal and insect addition")+
  theme_bw()+
  facet_grid(descriptor~site)

@@

@@

ggplot(fidata, aes(x = x, y = y)) +
  geom_point(size = 2, alpha = 0.4)+
  stat_smooth(data = fidata %>%
                filter(sign == "signif"),
              method ="lm")+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  ylab("Effect size of insects")+
  xlab("Effect size of fungi")+
  theme_bw()+
  facet_grid(descriptor~site)

@@

    print(summary(mod))

@@

classifier_RF

@@

# Variable importance plot
varImpPlot(classifier_RF)
```

@@

Explore that visually

@@

# Create data-set for a given treatment
deltaData <- data.frame()

# Change total bio variable into numeric
lgsData$tot_bio <- as.numeric(lgsData$tot_bio)

# Loop through each species in each garden and site and calculate their deltaBio
for (st in unique(lgsData$site)){

  for (gard in unique(lgsData$garden)){

    stGardDat <- lgsData %>%
      filter(site == st,
             garden == gard)

    if (dim(stGardDat)[1] == 0){
      # If subset is empty skip to another one
      next
    }

    if (length(unique(stGardDat$treatment)) < 2){
      # If there is species in only one treatment skip to another one
      next
    }

    specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
    specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)
    specInBoth <- specInCtr[specInCtr %in% specInTrt]

    cDat <- stGardDat %>%
      filter(unified_names %in% specInBoth,
             treatment == subsetVars[1]) %>%
      arrange(unified_names)
    tDat <- stGardDat %>%
      filter(unified_names %in% specInBoth,
             treatment == subsetVars[2]) %>%
      arrange(unified_names)

    # Calculate LRR for biomass and SLA change

    cDat$comparison <- subsetVars[2]

    if (subsetVars[2] == "h"){
      cDat$lrrtb <- log(tDat$tot_bio/cDat$tot_bio)
      cDat$lrrsla <- log(tDat$sla.d.m2kg1/cDat$sla.d.m2kg1)
    } else{
      cDat$lrrtb <- log(cDat$tot_bio/tDat$tot_bio)
      cDat$lrrsla <- log(cDat$sla.d.m2kg1/tDat$sla.d.m2kg1)
    }

    deltaData <- rbind(deltaData, cDat)

# Log transform estimated leaf area
deltaData <- deltaData %>%
          mutate(lest.leaf.area = log(est.leaf.area))%>%
  filter(!is.infinite(lest.leaf.area))

@@

# Log-response ratios of total biomass vs of sla

# DeltaBiomass vs SLA at each site
slaPlot <- ggplot(deltaData, aes(x = lsla.d, y = lrrtb, color = site)) +
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ x + I(x^2))

# There might be some effect of site.
# DeltaBiomass vs SLA at each site
wcPlot <- ggplot(deltaData, aes(x = water.cont, y = lrrtb, color = site)) +
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ x + I(x^2))
# Seems like the effect is not very strong either

herbPlot <- ggplot(deltaData, aes(x = perc.herb, y = lrrtb, color = site)) +
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ x + I(x^2)) + 
  scale_x_continuous(trans = 'log2')

ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)

```
It seems that in some cases the nonlinear relationship may be important. Generally water content looks linear for each elevation. We will include this and build also a model that models log-ratio biomass vs water content as a linear relationship. We arrive to the final definition of the full model.

```{r}
# Model explorations
simple.sla <- gls(
  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

just.sla <- gls(
  lrrtb ~ lsla.d*site + I(lsla.d^2)*site,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

# AIC(just.sla, simple.sla)
# anova(just.sla, simple.sla) # simple SLA is much bette
# tab_model(simple.sla) # But only the lrr SLA is significant.

# Run the full model.
full.sla <- gls(
  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + water.cont*site + I(water.cont^2)*site + perc.herb*site + I(perc.herb^2)*site + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

# Assuming linear relationships
full.sla.linear <- gls(
  lrrtb ~ lsla.d*site + water.cont*site + perc.herb*site + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

# anova(simple.sla, full.sla.linear) # full SLA linear is much better that the one using SLA
# tab_model(full.sla.linear)

full.sla.linear.noint <- gls(
  lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla + site,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

# anova(full.sla.linear, full.sla.linear.noint)
# Removing interactions does not improve the model significantly
# tab_model(full.sla.linear.noint)

# Add interaction only to the water content and sla
full.sla.linear.intslawc <- gls(
  lrrtb ~ lsla.d*site + water.cont*site + perc.herb + richness + labu + lrrsla + site,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

# anova(full.sla.linear.noint, full.sla.linear.intslawc)
# No effect
# tab_model(full.sla.linear.noint)

full.sla.linear.intsla <- gls(
  lrrtb ~ lsla.d*site + water.cont + perc.herb + richness + labu + lrrsla + site,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

# anova(full.sla.linear.noint, full.sla.linear.intsla)
# Any for of iteraction is not improving the model
tab_model(full.sla.linear.noint) # Thus the simplest model seems to be best.

# I further compare to the full.sla.linear.noint model.
# Only water content is modeled as linear
full.sla.custom <-  gls(
  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + water.cont*site + perc.herb*site + I(perc.herb^2)*site + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

# anova(full.sla.linear.noint, full.sla.custom)

full.sla.custom.noint <-  gls(
  lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + perc.herb + I(perc.herb^2) + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

full.sla.linear.noint.nosite <- gls(
  lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

# Withouth a site it is even better
# anova(full.sla.linear.noint, full.sla.linear.noint.nosite)

# The linear model remains the best one.
# plot(full.sla.linear.noint.nosite) # This plot looks best
# tab_model(full.sla.linear.noint.nosite)

@@

Table. Results of the selected model for the deltaBio.

```{r}
sjPlot::tab_model(full.sla.linear.noint.nosite)
```

@@

```{r, fig.cap="Observed vs predicted values for the lsls.d and "}
preddata <- data.frame(lsla.d = seq(from = min(deltaData$lsla.d, na.rm = T), 
                                  to = max(deltaData$lsla.d, na.rm = T), 
                                  length = 100),
                       water.cont = mean(deltaData$water.cont, na.rm = T), 
                       perc.herb= mean(deltaData$perc.herb, na.rm = T), 
                       richness= mean(deltaData$richness, na.rm = T), 
                       labu= mean(deltaData$labu, na.rm = T),
                       lrrsla= mean(deltaData$lrrsla, na.rm = T))

preddatalrrsla <- data.frame(lsla.d = mean(deltaData$lsla.d, na.rm = T),
                       water.cont = mean(deltaData$water.cont, na.rm = T), 
                       perc.herb= mean(deltaData$perc.herb, na.rm = T), 
                       richness= mean(deltaData$richness, na.rm = T), 
                       labu= mean(deltaData$labu, na.rm = T),
                       lrrsla=  seq(from = min(deltaData$lrrsla, na.rm = T), 
                                  to = max(deltaData$lrrsla, na.rm = T), 
                                  length = 100))

visData <- deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label)

predvals <- AICcmodavg::predictSE.gls(full.sla.linear.noint.nosite, newdata = preddata,
                    newdtype = 'response')

preddata$predvals <- predvals$fit
preddata$predse <- predvals$se.fit

predvalslrrsla <- AICcmodavg::predictSE.gls(full.sla.linear.noint.nosite, newdata = preddatalrrsla,
                    newdtype = 'response')

preddatalrrsla$predvals <- predvalslrrsla$fit
preddatalrrsla$predse <- predvalslrrsla$se.fit

# For lsla.d
p1 <- ggplot(preddata, aes(x = lsla.d, y = predvals)) + 
  geom_line()+
  geom_ribbon(aes(ymin = predvals - predse,
                  ymax = predvals + predse), 
              alpha = 0.1) + 
  geom_point(data = visData, aes(x = lsla.d, y = lrrtb)) + 
  ggtitle("Log(SLA) vs Log-response ratio of a species' biomass")

# For lrrsla
p2 <- ggplot(preddatalrrsla, aes(x = lrrsla, y = predvals)) + 
  geom_line()+
  geom_ribbon(aes(ymin = predvals - predse,
                  ymax = predvals + predse), 
              alpha = 0.1) + 
  geom_point(data = visData, aes(x = lrrsla, y = lrrtb)) +
  ggtitle("Delta(SLA) vs Log-response ratio of a species' biomass")

# ggplot(visData, aes(x = lsla.d, y = lrrtb)) + 
#   geom_point(aes(size = tot_bio, alpha = 0.05), pch = 19)+
#   geom_point(aes(x = lsla.d, y = predvals, col = "red")) + 
#   geom_smooth(aes(x = lsla.d, y = predvals), method = "lm")

ggpubr::ggarrange(p1,p2)

@@

```{r, include=FALSE}
# The Focused model
# Only the predictive power of SLA, water content and percentage herbivory on deltaBio

# Wanang ----
wdata <- deltaData %>% 
  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
  filter(site == "wanang")

# dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d,
#                      data = wdata,
#                      method = "ML",
#                      correlation = corMartins(2.718282, 
#                                               phy = myTree.rand.poly, 
#                                               form = ~unified_names,
#                                               fixed = T),
#                      na.action = na.omit)
dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
                     data = wdata,
                     method = "ML",
                     correlation = corMartins(2.718282, 
                                              phy = myTree.rand.poly, 
                                              form = ~unified_names,
                                              fixed = T),
                     na.action = na.omit)

# Model results

# Predict values
pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.w, wdata, 
                                       se.fit = T)
wdata <- wdata %>%
  mutate(preds = pred.vals$fit,
         se.preds = pred.vals$se.fit,
         se.upr = preds + se.preds,
         se.lwr = preds - se.preds)

# print the MSE
mse_focused_df <- data.frame(mse = with(wdata, mean((lrrtb - preds))^2),
                             treatment = "Insecticide", site = "Wanang")

# Numba ----
ndata <- deltaData %>% 
  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
  filter(site == "numba")

# dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d,
#                      data = ndata,
#                      method = "ML",
#                      correlation = corMartins(2.718282, 
#                                               phy = myTree.rand.poly, 
#                                               form = ~unified_names,
#                                               fixed = T),
#                      na.action = na.omit)
dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
                     data = ndata,
                     method = "ML",
                     correlation = corMartins(2.718282, 
                                              phy = myTree.rand.poly, 
                                              form = ~unified_names,
                                              fixed = T),
                     na.action = na.omit)

# Predict values
pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.n, ndata, 
                                       se.fit = T)
ndata <- ndata %>%
  mutate(preds = pred.vals$fit,
         se.preds = pred.vals$se.fit,
         se.upr = preds + se.preds,
         se.lwr = preds - se.preds)

# print the MSE
mse_focused_df <- rbind(mse_focused_df,  
                        data.frame(mse = with(ndata, mean((lrrtb - preds))^2),
                                   treatment = "Insecticide", site = "Numba"))

# Yawan ----
ydata <- deltaData %>% 
  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
  filter(site == "yawan")

# The model
# dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d,
#                      data = ydata,
#                      method = "ML",
#                      correlation = corMartins(2.718282, 
#                                               phy = myTree.rand.poly, 
#                                               form = ~unified_names,
#                                               fixed = T),
#                      na.action = na.omit)

dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
                     data = ydata,
                     method = "ML",
                     correlation = corMartins(2.718282,
                                              phy = myTree.rand.poly,
                                              form = ~unified_names,
                                              fixed = T),
                     na.action = na.omit)

# Predict values
pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.y, ydata, 
                                       se.fit = T)
ydata <- ydata %>%
  mutate(preds = pred.vals$fit,
         se.preds = pred.vals$se.fit,
         se.upr = preds + se.preds,
         se.lwr = preds - se.preds)

# print the MSE
mse_focused_df <- rbind(mse_focused_df,  
                        data.frame(mse = with(ydata, mean((lrrtb - preds))^2),
                                   treatment = "Insecticide", site = "Yawan"))

library(sjPlot)

tab_model(dd.gls.focsed.w, dd.gls.focsed.n, dd.gls.focsed.y)

# fig.cap = "Empirical and predicted values of a simple model of biomass change vs logarithm of SLA for each studied sites."}

# Plot the relationship and add predicted values
wpredPlot <- ggplot(wdata, aes(y = lrrtb, x = lsla.d))+
  geom_point() +
  geom_line(data = wdata, aes(x = lsla.d, y = preds), lwd = 1,
            lty = 2)+
  # geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) +
  ggtitle("Wanang 200 m a.s.l.")

npredPlot <- ggplot(ndata, aes(y = lrrtb, x = lsla.d))+
  geom_point() +
  geom_line(data = ndata, aes(x = lsla.d, y = preds), 
            lwd = 1, lty = 2)+
  # geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) + 
  ggtitle("Numba 750 m a.s.l.")

ypredPlot <- ggplot(ydata, aes(y = lrrtb, x = lsla.d))+
  geom_point() +
  geom_line(data = ydata, aes(x = lsla.d, y = preds), lwd = 1)+
  geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) + 
  ggtitle("Yawan 1900 m a.s.l.")

ggpubr::ggarrange(wpredPlot, npredPlot, ypredPlot,
                  nrow = 1)
# Plot the relationship and add predicted values
ggplot(rbind(ydata, ndata, wdata), aes(y = lrrtb, x = lsla.d, color = site))+
  geom_point() + 
  geom_line(data = rbind(ydata, ndata, wdata), 
            aes(x = lsla.d, y = preds), lwd = 1)+
  geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15)

sjPlot::tab_df(mse_focused_df)

@@

The above shows that focused model, where we used SLA as the only predictor of the plant change in biomass has a declining ability to predict change in biomass at the insecticide plot.

# Fungicide

```{r}
subsetVars <- c("c", "f")

ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)

@@

source('code/model_definitions.R')

# Model selection.
# 1. Full model (quadratic, and site interactions) vs no traits
# anova(no.traits, full.model)
# anova(no.traits, full.model.noint)
# anova(no.traits, all.linear)
# anova(no.traits, all.linear.noint)

# 2. Individual traits

# SLA
# Check quadratic
# anova(no.traits, just.sla.quadratic.noint)
# Check linear
# anova(no.traits, just.sla.linear.noint)
# Check quadratic interation
# anova(no.traits, just.sla.quadratic)
# Check linear interaction
# anova(no.traits, just.sla.linear)

# WATER
# Check quadratic
# anova(no.traits, just.water.quadratic.noint)
# Check linear
# anova(no.traits, just.water.linear.noint)
# Check quadratic interation
# anova(no.traits, just.water.quadratic)
# Check linear interaction
# anova(no.traits, just.water.linear)

# HERBIVORY
# Check quadratic
# anova(no.traits, just.herb.quadratic.noint)
# Check linear
# anova(no.traits, just.herb.linear.noint)
# Check quadratic interation
# anova(no.traits, just.herb.quadratic)
# Check linear interaction
# anova(no.traits, just.herb.linear)

# 2. Model with all traits linear/quadratic, what combination?

# quadratic SLA, HERB
# anova(no.traits, q.sla.q.herb)
# anova(no.traits, q.sla.q.herb.noint)
# quadratic SLA, linear HERB
# anova(no.traits, q.sla.l.herb)
# anova(no.traits, q.sla.l.herb.noint)
# linear SLA, quadratic HERB
# anova(no.traits, l.sla.q.herb)
# anova(no.traits, l.sla.q.herb.noint)
# linear SLA, linear HERB
# anova(no.traits, l.sla.l.herb)
# anova(no.traits, l.sla.l.herb.noint)

# quadratic HERB, WATER
# anova(no.traits, q.herb.q.water)
# anova(no.traits, q.herb.q.water.noint)
# quadratic HERB, linear WATER
# anova(no.traits, q.herb.l.water)
# anova(no.traits, q.herb.l.water.noint)
# linear HERB, quadratic WATER
# anova(no.traits, l.herb.q.water)
# anova(no.traits, l.herb.q.water.noint)
# linear HERB, linear WATER
# anova(no.traits, l.herb.l.water)
# anova(no.traits, l.herb.l.water.noint)

# quadratic SLA, WATER
# anova(no.traits, q.sla.q.water)
# anova(no.traits, q.sla.q.water.noint)
# quadratic SLA, linear WATER
# anova(no.traits, q.sla.l.water)
# anova(no.traits, q.sla.l.water.noint)
# linear SLA, quadratic WATER
# anova(no.traits, l.sla.q.water)
# anova(no.traits, l.sla.q.water.noint)
# linear SLA, linear WATER
# anova(no.traits, l.sla.l.water)
# anova(no.traits, l.sla.l.water.noint)

# So the best model is the one without the traits!
tab_model(no.traits)

# Additional herbivores

@@

ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)

@@

source('code/model_definitions.R')

@@

# Model selection.
# 1. Full model (quadratic, and site interactions) vs no traits
anova(no.traits, full.model) # full model is much better
anova(full.model, full.model.noint)

# interaction model is as good as the no interaction. Keep the model without the interaction

# 2. Is the quadratic relationship significant
# anova(full.model.noint, all.linear)            # AIC is lower for the full model noint
# anova(full.model.noint, all.linear.noint)      # AIC is lower for the full model noint

# Traits have nonlinear relationships with the delta biomass.

# 3. REMOVE SLA
# 3.1 anova(full.model.noint, q.herb.q.water.noint) # withouth sla is slightly better (AIC)
# 3.2 anova(full.model.noint, q.herb.q.water)
# 3.3 anova(full.model.noint, q.herb.l.water.noint) # Full model is than this !!!
# 3.4 anova(full.model.noint, q.herb.l.water)
# 3.5 anova(full.model.noint, l.herb.l.water.noint) # Full model still better
# 3.6 anova(full.model.noint, l.herb.l.water) #

# Including SLA did not mprove the model significantly
# tab_model(q.herb.q.water.noint)

# But if we were to start first by remving perc.herbivory?
# anova(full.model.noint, q.sla.q.water.noint) # Model without herbivory is better
# anova(full.model.noint, q.sla.q.water)       # Full model is better
# anova(q.sla.q.water, q.sla.q.water.noint)    # No interaction is better
# anova(full.model.noint, q.sla.l.water.noint)   # Full model is still better
# anova(full.model.noint, q.sla.l.water)         # Full model wins again
# anova(full.model.noint, l.sla.l.water.noint)   # And again by AIC criterium
# anova(full.model.noint, l.sla.l.water)         # And again

# But if we were to start first by remving water.content?
# anova(full.model.noint, q.sla.q.herb.noint) # Model with herbivory is much better
# anova(full.model.noint, q.sla.q.herb)       # Full model is better
# anova(q.sla.q.herb, q.sla.q.herb.noint)    # No interaction is better
# anova(full.model.noint, q.sla.l.herb.noint)   # Full model is still better
# anova(full.model.noint, q.sla.l.herb)         # Full model wins again
# anova(full.model.noint, l.sla.l.herb.noint)   # And again by AIC criterium
# anova(full.model.noint, l.sla.l.herb)         # And again

# The above suggests that models without hebivory and sla do better job. Water content is the key trait that helps predict the change in biomass in response to the treatment.

# SLA
# Check quadratic
# anova(full.model.noint, just.sla.quadratic.noint)
# Check linear
# anova(full.model.noint, just.sla.linear.noint)
# Check quadratic interation
# anova(full.model.noint, just.sla.quadratic)
# Check linear interaction
# anova(full.model.noint, just.sla.linear)

# WATER
# Check quadratic
anova(full.model.noint, just.water.quadratic.noint) # Model with just water content

# Is the quadratic form necessary
anova(just.water.quadratic.noint, just.water.linear.noint) # no difference but quadratic has lower AIC

# So the best model is the one without the traits!
tab_model(just.water.quadratic.noint)

@@

>>>>

```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
# The source file loads the data preparation script, where the raw data is read and summarised for further analyses
source('code/log_ratio_extraction.R')

library(ggplot2)
library(dplyr)
library(sjPlot)
library(AICcmodavg)

@@

Figure 1. Median, 1st, second quadrilles, 1.5 IRQ distance, and outliers (labeled) for logarithm of biomass, richness diversity, and density for the three studied location/elevation.
```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure S1 Characteistics"}

@@

# summary(trtlm)
# summary(trtgls)

@@

Table S2. Two models testing difference of LRR of predator exclosures from zero at each elevation.
```{r}
sjPlot::tab_df(model_comp)
```

@@

      # np.test.p <- cor.test(lrrtd$x, lrrtd$y, method = c('pearson'))$p.value
      # Test for noramlity
      # print(shapiro.test(lrrtd$x))
      # print(shapiro.test(lrrtd$y))

@@

      # print(lm.test.p)

@@

# ggplot(cpdata, aes(x = x, y = y)) +
#   geom_point(size = 2, alpha = 0.4)+
#   stat_smooth(data = cpdata %>%
#                 filter(sign == "signif"),
#               method ="lm")+
#   geom_hline(yintercept = 0)+
#   geom_vline(xintercept = 0)+
#   ylab("Effect size of insects")+
#   xlab("Effect size")+
#   theme_bw()+
#   facet_grid(response*descriptor~site)

@@

# ggplot(fidata, aes(x = x, y = y)) +
#   geom_point(size = 2, alpha = 0.4)+
#   stat_smooth(data = fidata %>%
#                 filter(sign == "signif"),
#               method ="lm")+
#   geom_hline(yintercept = 0)+
#   geom_vline(xintercept = 0)+
#   xlab("Predator removal")+
#   ylab("Predator removal and insect addition")+
#   theme_bw()+
#   facet_grid(descriptor~site)

@@

      # Test for normality
      # print(shapiro.test(lrrtd$x))
      # print(shapiro.test(lrrtd$y))

@@

# ggplot(fidata, aes(x = x, y = y)) +
#   geom_point(size = 2, alpha = 0.4)+
#   stat_smooth(data = fidata %>%
#                 filter(sign == "signif"),
#               method ="lm")+
#   geom_hline(yintercept = 0)+
#   geom_vline(xintercept = 0)+
#   ylab("Effect size of insects")+
#   xlab("Effect size of fungi")+
#   theme_bw()+
#   facet_grid(descriptor~site)

@@

    # print(summary(mod))

@@

# classifier_RF

@@

# Variable importance plot
# varImpPlot(classifier_RF)
```

@@

It seems that in some cases the nonlinear relationship may be important. Generally water content looks linear for each elevation. We will include this and build also a model that models log-ratio biomass vs water content as a linear relationship. We arrive to the final definition of the full model.

@@

# ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)

@@

source('code/model_definitions_pgls.R')

# Model selection.

# 1. Full model (quadratic, and site interactions) vs no traits
# anova(no.traits, full.model)            # full model is better

# 2. Removal of individual traits

# 2.1 Remove SLA
# anova(full.model, q.herb.q.water) # no-difference, keep model without the SLA a simpler model?
# 2.2 Remove herbivory
# anova(full.model, q.sla.q.water.noint) # Model without herbivory is significantly worse
# 2.3 Remove water content?
# anova(full.model, q.sla.q.herb) # Model without water is significantly worse

# Herbivory and water are the most important!!!

# 3. Reduce model even further - 1 trait models.
# anova(q.herb.q.water, just.sla.quadratic)   # Two trait model better
# anova(q.herb.q.water, just.herb.quadratic)  # Two trait model better
# anova(q.herb.q.water, just.water.quadratic) # Two trait model better

# 4. Quadratic vs linear
# anova(q.herb.q.water, l.herb.l.water) # Quadratic model is still better

# 5. Interaction reduction
# anova(q.herb.q.water, q.herb.q.water.noint) # Model with interaction is better
# anova(q.herb.q.water, just.water.quadratic.noint) # Model with interaction is better
# anova(q.herb.q.water, just.water.linear.noint)

@@

```{r, results='asis'}
# So the best model is the one without the traits!
# tab_model(q.herb.q.water)
# tab_model(just.water.quadratic.noint)
i.mod <- q.herb.q.water
i.dD <- deltaData
```

@@

@@

#### Fungicide

subsetVars <- c("c", "f")

# ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)

@@

source('code/model_definitions_pgls.R')

# Model selection.

# 1. Full model (quadratic, and site interactions) vs no traits
# anova(no.traits, full.model)            # full model is better

# 2. Removal of individual traits

# 2.1 Remove SLA
# anova(full.model, q.herb.q.water) # without sla is slightly better (AIC)

# 2.2 Remove herbivory
# anova(full.model, q.sla.q.water) # Model without herbivory is not better

# Both q.herb.q.water and q.sla.q.water are simpler, but which one is better?
# anova(full.model, q.sla.q.herb) # Model without water is significantly worse

# 2.3 Remove water content?
# anova(full.model, q.sla.q.herb) # Model with herbivory is much better!
# Water is important!!!

# anova(q.herb.q.water, q.sla.q.water) # q.sla.q.water is better based on AIC

# 3. Reduce model even further - 1 trait models.
# anova(q.sla.q.water, just.sla.quadratic)  # q.sla.q.water is still better
# anova(q.sla.q.water, just.herb.quadratic) # q.sla.q.water is still better
# anova(q.sla.q.water, just.water.quadratic) # model with just water is as good as q.sla.q.water

# model with just water is not different form the full model, and it has lower AIC

# Seems like they all stay!

# 4. Quadratic vs linear
# anova(just.water.quadratic, just.water.linear) # Linear is simpler model still better

# 5. Interaction reduction
# anova(just.water.linear, just.water.linear.noint) # Interaction does not bring anything to the model

@@

# So the best model is the one with only water!
# tab_model(just.water.linear.noint)
f.mod <- just.water.linear.noint
f.dD <- deltaData
```

#### Additional herbivores

@@

# ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)

@@

source('code/model_definitions_pgls.R')

@@

# Model selection.

# 1. Full model (quadratic, and site interactions) vs no traits
# anova(no.traits, full.model)            # full model is better

# 2. Removal of individual traits

# 2.1 Remove SLA
# anova(full.model, q.herb.q.water) # without sla is slightly better (AIC)

# 2.2 Remove herbivory
# anova(full.model, q.sla.q.water) # Model without herbivory is not better

# 2.3 Remove water content?
# anova(full.model, q.sla.q.herb) # Model with herbivory is much better

# Two candidate models:
# anova(q.herb.q.water,q.sla.q.water) # q.sla.q.water is a bit better (AIC)

# 3. Reduce model even further - 1 trait models.
# anova(q.sla.q.water, just.sla.quadratic) # MOdel with water is significantly better than just sla
# anova(q.sla.q.water, just.herb.quadratic) # It is better than just herbivory
# anova(q.sla.q.water, just.water.quadratic) # But is not better than model with just water

# Seems like they all stay!

# 4. Quadratic vs linear
# anova(just.water.quadratic, just.water.linear) # Linear model is not simpler, it has slightly higher AIC

# 5. Interaction reduction
# anova(just.water.linear, just.water.linear.noint) # Interaction model is significantly better
```

```{r, results='asis'}
# So the best model is ... !
# tab_model(just.water.linear)
h.mod <- just.water.linear
h.dD <- deltaData
```

#### Predators

```{r}
subsetVars <- c("c", "p")

# ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)

@@

source('code/model_definitions_pgls.R')

# Model selection.

# 1. Full model (quadratic, and site interactions) vs no traits
# anova(no.traits, full.model)            # full model is better

# 2. Removal of individual traits

# 2.1 Remove SLA
# anova(full.model, q.herb.q.water) # Full model is significanlty better

# 2.2 Remove herbivory
# anova(full.model, q.sla.q.water.noint) # Full model still better

# 2.3 Remove water content?
# anova(full.model, q.sla.q.herb) # Full model still better

# 3. Reduce model even further - 1 trait models.
# anova(full.model, just.sla.quadratic)
# anova(full.model, just.herb.quadratic)
# anova(full.model, just.water.quadratic)

# Seems like they all stay!

# 4. Quadratic vs linear
# anova(full.model, all.linear) # Full model still better

# 5. Interaction reduction
# anova(full.model, full.model.noint) # Full model still better
```

```{r, results='asis'}
# So the best model is the one without the traits!
# tab_model(full.model)
p.mod <- full.model
p.dD <- deltaData
```

### SLA models for treatments

tab_model(p.mod, i.mod, h.mod, f.mod,
          pred.labels = c("Intercept", 
                          "Log[SLA]-linear", 
                          "Numba", 
                          "Yawan",
                          "Log[SLA]-quadratic", 
                          "Water content - linear", 
                          "Water content - quadratic",
                          "Herbivory [%] - linear",
                          "Herbivory [%] - quadratic",
                          "Species richness",
                          "Abundance",
                          "LRR SLA",
                          "Log[SLA]-linear : Numba",
                          "Log[SLA]-linear : Yawan",
                          "Log[SLA]-quadratic : Numba",
                          "Log[SLA]-quadratic : Yawan",
                          "Water content-linear : Numba",
                          "Water content-linear : Yawan",
                          "Water content-quadratic : Numba",
                          "Water content-quadratic : Yawan",
                          "Herbivory-linear : Numba",
                          "Herbivory-linear : Yawan",
                          "Herbivory-quadratic : Numba",
                          "Herbivory-quadratic : Yawan"),
          dv.labels = c("Predator", "Insecticide", "Extra herbivory", "Fungicide"),
          string.pred = "Coeffcient",
          string.ci = "Conf. Int (95%)",
          string.p = "P-Value")

### Predictions and visualizations

```{r}
source("code/predictAndVisualize_function.R")

# Predator --------------------------
pslap <- predictAndVisualize(p.mod, 
                    p.dD,
                    "lsla.d",
                    n = 100,
                    titlestring = "Predator - A",
                    ylabstring = "Log-response ratio of total biomass.",
                    xlabstring = "Log(SLA)")

# Focal traits

# Co-variates
psitep <- predictAndVisualize(p.mod, 
                    p.dD,
                    "fsite",
                    n = 100,
                    titlestring = "Predator - B",
                    ylabstring = "Log-response ratio of total biomass.",
                    xlabstring = "Sites")

plabup <- predictAndVisualize(p.mod, 
                    p.dD,
                    "labu",
                    n = 100,
                    titlestring = "Predator - C",
                    ylabstring = "Log-response ratio of total biomass.",
                    xlabstring = "Log(stem number)")

# Insecticide-----------------------
ilabup <- predictAndVisualize(i.mod, 
                    i.dD,
                    "labu",
                    n = 100,
                    titlestring = "Insecticide - C",
                    ylabstring = "Log-response ratio of total biomass.",
                    xlabstring = "Log(stem number)")

# Extra herbivory-------------------
hsitep <- predictAndVisualize(h.mod, 
                    h.dD,
                    "fsite",
                    n = 100,
                    titlestring = "Herbivory - A",
                    ylabstring = "Log-response ratio of total biomass.",
                    xlabstring = "Sites")
